### 2024-03-15 ms.qmd
Updates

Behavioural flexibility describes the ability of individuals to adjust behaviour adaptively in response to internal or external changes in the environment [@brown2010behavioral]. It comprises a suite of behavioural aspects like problem-solving and innovation [@brown2010behavioral], that are predicted to be crucial for invasion success [@wright2010behavioral; @chapple2012can; @szabo2020can]. Species with greater behavioural flexibility are expected to be more likely to survive and establish in new habitats because behaving flexibly enhances their capacity to exploit new resources and face new challenges [@wright2010behavioral; @chapple2012can; @szabo2020can]. For instance, in closely related bird species successful invaders show a higher frequency of foraging innovations [@sol2000behavioural]. According to the 'adaptive flexibility hypothesis', behavioural flexibility is most relevant during introduction, establishment, and spread of an invasive population [@wright2010behavioral; @chapple2012can; @szabo2020can]. In mammals, birds, reptiles and amphibians, relative brain size has been shown to be a good predictor of behavioural flexibility and invasive success [@sol2000behavioural; @sol2002behavioural; @sol2008brain; @amiel_smart_2011]. However, the brain is also sensitive to environmental changes during the early stages of development [@zhu_prenatal_2004] creating uncertainty about how environmental insults faced by early invaders might impact behavioural flexibility and thus the probability of invasion success.    
**BETTER LINK**In this scenario, prenatal glucocorticoids (GCs) and the thermal environment can be salient factors in the study of behavioural flexibility within invasion biology. GCs are the main group of hormones related to animals’ response towards alterations in homeostasis [@sapolsky_how_2000] that may become relevant when facing new environments and their new thermal conditions [@Crino_2023]. In addition, there is evidence that developmental GCs and thermal environment can impact cognition in different taxa [@jonson1976effect; @zhu_prenatal_2004; @amiel_hotter_2012; @lui2017chronic]. For instance, young Sprague-Dawley rats treated with corticosterone (CORT) - the principal GC in rodents, birds, and reptiles [@Crino_2023] - performed worse than controls in a reversal learning task [@lui2017chronic], while cold incubation temperatures in the eastern three-lined skink resulted in significant impairments with associative and reversal learning [@clark_colour_2014]. Moreover, the cognitive effects of both factors are triggered by similar physiological mechanisms intertwined with brain development [@jonson1976effect; @coomber_independent_1997; @zhu_prenatal_2004; @amiel_effects_2017], an aspect considered essential in successful invaders [@sol2000behavioural; @sol2002behavioural; @sol2008brain; @amiel_smart_2011]. 
Here, we studied the interactive effects of prenatal temperature and CORT on behavioural flexibility in two species with distinct histories of invasiveness, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). _L. delicata_ has been succesful in colonising different areas around the globe [@baker1979rainbow; @chapple2013divergent; @chapple2015deliinvLHI], but the same is not true of _L. guichenoti_ despite the two species having similar life-histories and overlapping distributions [@chapple_know_2011; @chapple_biology_2014]. Personality has been shown to be different between these two species [@chapple_know_2011; @chapple2012can; @bezzina2014does], which could be related to general cognitive differences that are important for invasion success. Additionally, early environments are known to affect suites of morphological and behavioural traits differently in each species [**REFS**], but the interactive role of maternal CORT and early temperatures have not yet been explored. Differences in the relative impact that early environments have on behavioural flexibility may help explain differing invasion success between the two species. We predicted that _L. delicata_ would, on average, perform more proficiently in the task compared to _L. guichenoti_. We also anticipated prenatal CORT and temperature to affect the behavioural flexibility of both species, and that _L. guichenoti_ would be more impacted by the prenatal environment than _L. delicata_.

@@

Both species learnt across all treatments (@fig-results B, D), but we did not find any difference between treatments or species in the rate of learning. For _L. delicata_, hormone treatment (Control - CORT = `r format_dec(mean(c(deli_ControlCold, deli_ControlHot)) - mean(c(deli_CORTCold + deli_CORTHot)), 3)`, p~mcmc~ = `r format_p(pmcmc(c(deli_ControlCold, deli_ControlHot) - c(deli_CORTCold, deli_CORTHot)), 3)`), incubation temperature (Hot - Cold = `r format_dec(mean(c(deli_CORTHot, deli_ControlHot)) - mean(c(deli_CORTCold, deli_ControlCold)),3)`, p~mcmc~ = `r format_p(pmcmc(c(deli_CORTHot, deli_ControlHot) - c(deli_CORTCold, deli_ControlCold)),3)`), or the interaction ([(Control-Hot - CORT-Hot) - (Control-Cold - CORT-Cold)] = `r format_dec(mean(deli_ControlHot - deli_CORTHot) - mean(deli_ControlCold - deli_CORTCold), 3)`, p~mcmc~ = `r format_p(pmcmc(c(deli_ControlHot, deli_CORTHot) - c(deli_ControlCold, deli_CORTCold)), 3)`) did not affect learning (see @fig-results A, B). We similarly found no significant effects of  hormone (Control - CORT = `r format_dec(mean(c(guich_ControlCold, guich_ControlHot)) - mean(c(guich_CORTCold, guich_CORTHot)), 3)`, p~mcmc~ = `r format_p(pmcmc(c(guich_ControlCold, guich_ControlHot) - c(guich_CORTCold, guich_CORTHot)), 3)`), incubation temperature (Hot - Cold = `r format_dec(mean(c(guich_CORTHot, guich_ControlHot)) - mean(c(guich_CORTCold, guich_ControlCold)),3)`, p~mcmc~ = `r format_p(pmcmc(c(guich_CORTHot, guich_ControlHot) - c(guich_CORTCold, guich_ControlCold)),3)`), **or the interaction ([(Control-Hot - CORT-Hot) - (Control-Cold - CORT-Cold)] = `r format_dec(mean(guich_ControlHot - guich_CORTHot) - mean(guich_ControlCold - guich_CORTCold), 3)`, p~mcmc~ = `r format_p(pmcmc(mean(guich_ControlHot - guich_CORTHot) - mean(guich_ControlCold - guich_CORTCold)), 3)`) on learning in _L. guichenoti_ (see @fig-results C, D).**

>>>>

Behavioural flexibility describes the ability of individuals to adjust behaviour adaptively in response to internal or external changes in the environment [@brown2010behavioral]. It comprises a suite of behavioural aspects like problem-solving and innovation [@brown2010behavioral] predicted to be crucial for invasion success [@wright2010behavioral; @chapple2012can; @szabo2020can]. Species with greater behavioural flexibility are expected to be more likely to survive and establish in new habitats because behaving flexibly enhances their capacity to exploit new resources and face new challenges [@wright2010behavioral; @chapple2012can; @szabo2020can]. For instance, in closely related bird species successful invaders show a higher frequency of foraging innovations [@sol2000behavioural]. According to the 'adaptive flexibility hypothesis', behavioural flexibility is most relevant during introduction, establishment, and spread of an invasive population [@wright2010behavioral; @chapple2012can; @szabo2020can]. In mammals, birds, reptiles and amphibians, relative brain size has been shown to be a good predictor of behavioural flexibility and invasive success [@sol2000behavioural; @sol2002behavioural; @sol2008brain; @amiel_smart_2011]. However, the brain is also sensitive to environmental changes during the early stages of development [@zhu_prenatal_2004] creating uncertainty about how environmental insults faced by early invaders might impact behavioural flexibility and thus the probability of invasion success.  
New environments faced by invasive populations introduce organisms to a plethora of unfamiliar conditions that can entail alterations in homeostasis. This disruption triggers a suite of physiological and behavioural responses that are mediated by glucocorticoids (GCs), the "stress"-related hormones [@sapolsky_how_2000]. These GCs can be transmitted to the offspring and  influence its phenotype through transgenerational effects [@Crino_2023]. At the same time, other aspects of the environment, such as temperature, can impact the offspring phenotype through direct effects [@noble_developmental_2018]. As such, GCs and  temperature are predicted to interact to influence the offspring's phenotype. In this sense, both factors are known to affect similar aspects of brain development [@jonson1976effect; @coomber_independent_1997; @zhu_prenatal_2004; @amiel_effects_2017] and also cognitive abilities such as reversal learning [@clark_colour_2014; @lui2017chronic]. However, the interactive repercussions of prenatal GCs and temperature on behavioural flexibility have not been explored.  
Here, we studied the effects of prenatal temperature and CORT, the main GC in reptiles [@Crino_2023], on behavioural flexibility in two species with distinct histories of invasiveness, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). _L. delicata_ has been succesful in colonising different areas around the globe [@baker1979rainbow; @chapple2013divergent; @chapple2015deliinvLHI], but the same is not true of _L. guichenoti_ despite the two species having similar life-histories and overlapping distributions [@chapple_know_2011; @chapple_biology_2014]. Personality has been shown to be different between these two species [@chapple_know_2011; @chapple2012can; @bezzina2014does], which could be related to general cognitive differences that are important for invasion success. Additionally, early environments are known to affect suites of morphological and behavioural traits differently in each species [@carrasco2023maternal; @kar2023heritability], but the interactive role of maternal CORT and early temperatures have not yet been explored. Differences in the relative impact that early environments have on behavioural flexibility may help explain differing invasion success between the two species. We predicted that _L. delicata_ would, on average, perform more proficiently in the task compared to _L. guichenoti_. We also anticipated prenatal CORT and temperature to affect the behavioural flexibility of both species, and that _L. guichenoti_ would be more impacted by the prenatal environment than _L. delicata_.

@@

Both species learnt across all treatments (@fig-results B, D), but we did not find any difference between treatments or species in the rate of learning. For _L. delicata_, hormone treatment (Control - CORT = `r format_dec(mean(c(deli_ControlCold, deli_ControlHot)) - mean(c(deli_CORTCold + deli_CORTHot)), 3)`, p~mcmc~ = `r format_p(pmcmc(c(deli_ControlCold, deli_ControlHot) - c(deli_CORTCold, deli_CORTHot)), 3)`), incubation temperature (Hot - Cold = `r format_dec(mean(c(deli_CORTHot, deli_ControlHot)) - mean(c(deli_CORTCold, deli_ControlCold)),3)`, p~mcmc~ = `r format_p(pmcmc(c(deli_CORTHot, deli_ControlHot) - c(deli_CORTCold, deli_ControlCold)),3)`), or the interaction ([(Control-Hot - CORT-Hot) - (Control-Cold - CORT-Cold)] = `r format_dec(mean(deli_ControlHot - deli_CORTHot) - mean(deli_ControlCold - deli_CORTCold), 3)`, p~mcmc~ = `r format_p(pmcmc(c(deli_ControlHot, deli_CORTHot) - c(deli_ControlCold, deli_CORTCold)), 3)`) did not affect learning (see @fig-results A, B). We similarly found no significant effects of  hormone (Control - CORT = `r format_dec(mean(c(guich_ControlCold, guich_ControlHot)) - mean(c(guich_CORTCold, guich_CORTHot)), 3)`, p~mcmc~ = `r format_p(pmcmc(c(guich_ControlCold, guich_ControlHot) - c(guich_CORTCold, guich_CORTHot)), 3)`), incubation temperature (Hot - Cold = `r format_dec(mean(c(guich_CORTHot, guich_ControlHot)) - mean(c(guich_CORTCold, guich_ControlCold)),3)`, p~mcmc~ = `r format_p(pmcmc(c(guich_CORTHot, guich_ControlHot) - c(guich_CORTCold, guich_ControlCold)),3)`), **or the interaction ([(Control-Hot - CORT-Hot) - (Control-Cold - CORT-Cold)] = `r format_dec(mean(guich_ControlHot - guich_CORTHot) - mean(guich_ControlCold - guich_CORTCold), 3)`, p~mcmc~  `r format_p(pmcmc(mean(guich_ControlHot - guich_CORTHot) - mean(guich_ControlCold - guich_CORTCold)), 3)`) on learning in _L. guichenoti_ (see @fig-results C, D). (Control-Hot - CORT-Hot = `r format_dec(mean(guich_ControlHot) - mean(guich_CORTHot),3)`; p~mcmc~ = `r format_dec(pmcmc(guich_ControlHot - guich_CORTHot),3)`) or (Control-Cold - CORT-Cold = `r format_dec(mean(guich_ControlCold) - mean(guich_CORTCold),3)`; p~mcmc~ = `r format_dec(pmcmc(guich_ControlCold - guich_CORTCold),3)`)**
